---
title: "Training the Models"
author: "Nathan Byers"
date: "Sunday, April 12, 2015"
output:
  html_document:
    toc: yes
---

This page documents the model training process. For documentation on how the data was prepared, see [this page](http://rpubs.com/NateByers/machLearnDataPrep).


## Preprocessing

First, we library all of the necessary packages and load the training and test sets.

```{r, message=FALSE}
library(kernlab)
library(caret)
library(pROC)
# library the genefilter package (install from bioconductor first if it's not 
# installed already)
if(!require(genefilter)){
  source("http://bioconductor.org/biocLite.R")
  biocLite("genefilter")
  library(genefilter)
}
load("../Data/split_data.RData")
ls()
```

Then we make sure that the stage column in those data frames are factors.

```{r}
train_data$stage <- as.factor(train_data$stage)
test_data$stage <- as.factor(test_data$stage)
```

We'll do two kinds of preprocessing: 

1. Filter based on the intensity of the miRNA expression and the coefficient of variation, and
2. Filter based on the p-value from the t-test in the [data preparation stage](http://rpubs.com/NateByers/machLearnDataPrep) and then
transformation using principal components.

### Preprocessing: Method 1

Here we create a function that filters out miRNAs that do not have at least 20% of the sample 
expression value at 100 or greater. Then it filters out miRNAs that do not have coefficient of 
variation (sd/mean) between 0.7 and 10. (https://www.biostars.org/p/86981/)

```{r}
exprFilterFun <- filterfun(pOverA(p = 0.2, A = 100), cv(a = 0.7, b = 10))
```

Now we filter the training and test data sets.

```{r}
# make a matrix from the train_data data frame
train_matrix <- train_data[, -c(1:2)]
# transpose the matrix so that rows are miRNAs and columns are samples
train_matrix_t <- t(train_matrix)
# create a logical vector that indicates TRUE for features that meet the 
# filter requirments
expr_filter <- genefilter(train_matrix_t, exprFilterFun)
# see how many features are left
sum(expr_filter)
# subset the training data down to the features that meet the requirements
train_data_filtered <- train_data[, c(TRUE, TRUE, expr_filter)]
# subset test data
test_data_filtered <- test_data[, names(train_data_filtered)]
```

### Preprocessing: Method 2

For this method, we narrow the features down by choosing a p-value threshold, and then using
principal component analysis to transform the data into linear combinations of the miRNA expression data.

Here we choose a threshold of 0.1 for the p-values.
```{r}
sig_rna <- miRNA_pvalue[miRNA_pvalue$pvalue <= .1, "miRNA_id"]
train_data_sig <- data.frame(train_data[, 1:2], 
                             train_data[, sig_rna])
test_data_sig <- data.frame(test_data[, 1:2],
                            test_data[, sig_rna])
```

Then we apply PCA to the remaining data set. We'll choose a threshold of 95% (choose components
that account for 95% of the variation in the data).

```{r}
trans <- preProcess(train_data_sig[, -c(1:2)],
                   method = c("center", "scale", "pca"),
                   thresh = 0.95)
trans
```

117 components will be used

## Splitting

We'll be using the training samples, which totals 424 observations. To train the parameters for each model, we'll use two data splitting methods. 5- fold cross validation and 10 fold cross validation.

### 5-fold Cross Validation

Here we set the training control for 5-fold cross validation.

```{r}
ctrl5f <- trainControl(method = "cv", number = 5,
                      summaryFunction = twoClassSummary,
                      classProbs = TRUE)
```

### 10-fold Cross Validation

And here we set the training control for 10-fold cross validation.

```{r}
ctrl10f <- trainControl(method = "cv", number = 10,
                      summaryFunction = twoClassSummary,
                      classProbs = TRUE)
```


## Support Vector Machine (SVM)

The first algorithm is the support vector machine.

### SVM/PreProc1/5-foldCV

The first model we'll train is the support vector machine, using the first preproccessing method and 5-fold cross validation.

First we make a tuning grid that covers a range of sigma and cost values (Kuhn and Johnson 2013)
```{r} 
svmrGrid <- expand.grid(sigma = c(.00007, .00009, .0001, .0002),
                        C = 2^(-3:8))
```


Now we fit the model and report the run time. ROC is used to select the optimal model.

```{r, message=FALSE}
then <- Sys.time()
set.seed(477)
svm_pre1_5cv <- train(stage ~ .,
             data = train_data_filtered[, -1],
             method = "svmRadial",
             tuneGrid = svmrGrid,
             metric = "ROC",
             trControl = ctrl5f)
Sys.time() - then             
```

We predict using the test set and evaluate the performance.

```{r}
svm_pre1_5cv_pred <- predict(svm_pre1_5cv, 
                            newdata = test_data_filtered[, - c(1:2)])
svm_pre1_5cv_prob <- predict(svm_pre1_5cv, 
                            newdata = test_data_filtered[, - c(1:2)],
                            type = "prob")

svm_pre1_5cv.df <- data.frame(stage = test_data_filtered$stage, 
                             svm_pre1_5cv_pred, 
                             early_prob = svm_pre1_5cv_prob$early,
                             late_prob = svm_pre1_5cv_prob$late)

head(svm_pre1_5cv.df)

confusionMatrix(data = svm_pre1_5cv.df$svm_pre1_5cv_pred,
                reference = test_data_filtered$stage,
                positive = "late")

svm_pre1_5cv_rocCurve <- roc(response = test_data_filtered$stage,
                predictor = svm_pre1_5cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_filtered$stage)))

auc(svm_pre1_5cv_rocCurve)

ci.roc(svm_pre1_5cv_rocCurve)

plot(svm_pre1_5cv_rocCurve, legacy.axes = TRUE)

save(svm_pre1_5cv, svm_pre1_5cv_pred, svm_pre1_5cv_prob,
     file = "../Results/svm_pre1_5cv.RData")

```

### SVM/PreProc2/5-foldCV

SVM, preprocessing method 2, and 5 fold cross validation. (The default 
threshold for "pca" in the `preProc` argument is 95%)

```{r, message=FALSE}
then <- Sys.time()
set.seed(477)
svm_pre2_5cv <- train(stage ~ .,
             data = train_data_sig[, -1],
             method = "svmRadial",
             tuneGrid = svmrGrid,
             preProc = c("center", "scale", "pca"),
             metric = "ROC",
             trControl = ctrl5f)
Sys.time() - then

svm_pre2_5cv_pred <- predict(svm_pre2_5cv, 
                            newdata = test_data_sig[, - c(1:2)])
svm_pre2_5cv_prob <- predict(svm_pre2_5cv, 
                            newdata = test_data_sig[, - c(1:2)],
                            type = "prob")

svm_pre2_5cv.df <- data.frame(stage = test_data_sig$stage, 
                             svm_pre2_5cv_pred, 
                             early_prob = svm_pre2_5cv_prob$early,
                             late_prob = svm_pre2_5cv_prob$late)

head(svm_pre2_5cv.df)

confusionMatrix(data = svm_pre2_5cv.df$svm_pre2_5cv_pred,
                reference = test_data_sig$stage,
                positive = "late")

svm_pre2_5cv_rocCurve <- roc(response = test_data_sig$stage,
                predictor = svm_pre2_5cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_sig$stage)))

auc(svm_pre2_5cv_rocCurve)

ci.roc(svm_pre2_5cv_rocCurve)

plot(svm_pre2_5cv_rocCurve, legacy.axes = TRUE)

save(svm_pre2_5cv, svm_pre2_5cv_pred, svm_pre2_5cv_prob,
     file = "../Results/svm_pre2_5cv.RData")
```

### SVM/PreProc1/10-foldCV

SVM, preproccessing method 1, and 10 fold cross validation.


```{r, message=FALSE}
then <- Sys.time()
set.seed(477)
svm_pre1_10cv <- train(stage ~ .,
             data = train_data_filtered[, -1],
             method = "svmRadial",
             tuneGrid = svmrGrid,
             metric = "ROC",
             trControl = ctrl10f)
Sys.time() - then             
```

We predict using the test set and evaluate the performance.

```{r}
svm_pre1_10cv_pred <- predict(svm_pre1_10cv, 
                            newdata = test_data_filtered[, - c(1:2)])
svm_pre1_10cv_prob <- predict(svm_pre1_10cv, 
                            newdata = test_data_filtered[, - c(1:2)],
                            type = "prob")

svm_pre1_10cv.df <- data.frame(stage = test_data_filtered[, "stage"], 
                             svm_pre1_10cv_pred, 
                             early_prob = svm_pre1_10cv_prob$early,
                             late_prob = svm_pre1_10cv_prob$late)

head(svm_pre1_10cv.df)

confusionMatrix(data = svm_pre1_10cv.df$svm_pre1_10cv_pred,
                reference = test_data_filtered[, "stage"],
                positive = "late")

svm_pre1_10cv_rocCurve <- roc(response = test_data_filtered[, "stage"],
                predictor = svm_pre1_10cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_filtered[, "stage"])))

auc(svm_pre1_10cv_rocCurve)

ci.roc(svm_pre1_10cv_rocCurve)

plot(svm_pre1_10cv_rocCurve, legacy.axes = TRUE)

save(svm_pre1_10cv, svm_pre1_10cv_pred, svm_pre1_10cv_prob,
     file = "../Results/svm_pre1_10cv.RData")
```

### SVM/PreProc2/10-foldCV

SVM, preprocessing method 2, and 10-fold cross validation.

```{r, message=FALSE}
then <- Sys.time()
set.seed(477)
svm_pre2_10cv <- train(stage ~ .,
             data = train_data_sig[, -1],
             method = "svmRadial",
             tuneGrid = svmrGrid,
             preProc = c("center", "scale", "pca"),
             metric = "ROC",
             trControl = ctrl10f)
Sys.time() - then             

svm_pre2_10cv_pred <- predict(svm_pre2_10cv, 
                            newdata = test_data_sig[, - c(1:2)])
svm_pre2_10cv_prob <- predict(svm_pre2_10cv, 
                            newdata = test_data_sig[, - c(1:2)],
                            type = "prob")

svm_pre2_10cv.df <- data.frame(stage = test_data_sig[, "stage"], 
                             svm_pre2_10cv_pred, 
                             early_prob = svm_pre2_10cv_prob$early,
                             late_prob = svm_pre2_10cv_prob$late)

head(svm_pre2_10cv.df)

confusionMatrix(data = svm_pre2_10cv.df$svm_pre2_10cv_pred,
                reference = test_data_sig[, "stage"],
                positive = "late")

svm_pre2_10cv_rocCurve <- roc(response = test_data_sig[, "stage"],
                predictor = svm_pre2_10cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_sig[, "stage"])))

auc(svm_pre2_10cv_rocCurve)

ci.roc(svm_pre2_10cv_rocCurve)

plot(svm_pre2_10cv_rocCurve, legacy.axes = TRUE)

save(svm_pre2_10cv, svm_pre2_10cv_pred, svm_pre2_10cv_prob,
     file = "../Results/svm_pre2_10cv.RData")

```

## Random Forest (RF)

The second algorithm is random forest.

### RF/PreProc1/5-foldCV

Random forest, preprocessing method 1, 5 fold cross validation.

For random forest, we tune the mtry parameter using 5 values evenly spaced from 2
 to p, where p = # of predictors. We also use an ensemble of
 1000 trees (pg. 387, Kuhn and Johnson 2013)

```{r, message=FALSE}
p <- ncol(train_data_filtered) - 2
spaced.values <- round(seq(2, p, length.out = 7))
mtryValues <- spaced.values[2:6]

then <- Sys.time()
set.seed(477)
rf_pre1_5cv <- train(x = train_data_filtered[, -c(1:2)], 
               y = train_data_filtered$stage,
               method = "rf",
               ntree = 1000,
               tuneGrid = data.frame(mtry = mtryValues),
               importance = TRUE,
               metric = "ROC",
               trControl = ctrl5f)
Sys.time() - then

rf_pre1_5cv_pred <- predict(rf_pre1_5cv, 
                            newdata = test_data_filtered[, - c(1:2)])
rf_pre1_5cv_prob <- predict(rf_pre1_5cv, 
                            newdata = test_data_filtered[, - c(1:2)],
                            type = "prob")

rf_pre1_5cv.df <- data.frame(stage = test_data_filtered$stage, 
                             rf_pre1_5cv_pred, 
                             early_prob = rf_pre1_5cv_prob$early,
                             late_prob = rf_pre1_5cv_prob$late)

head(rf_pre1_5cv.df)

confusionMatrix(data = rf_pre1_5cv.df$rf_pre1_5cv_pred,
                reference = test_data_filtered$stage,
                positive = "late")

rf_pre1_5cv_rocCurve <- roc(response = test_data_filtered$stage,
                predictor = rf_pre1_5cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_filtered$stage)))

auc(rf_pre1_5cv_rocCurve)

ci.roc(rf_pre1_5cv_rocCurve)

plot(rf_pre1_5cv_rocCurve, legacy.axes = TRUE)

save(rf_pre1_5cv, rf_pre1_5cv_pred, rf_pre1_5cv_prob,
     file = "../Results/rf_pre1_5cv.RData")

```

### RF/PreProc2/5-foldCV

Random forest, preprocessing method 2, 5 fold cross validation.

```{r, message=FALSE}
then <- Sys.time()
set.seed(477)
rf_pre2_5cv <- train(x = train_data_sig[, -c(1:2)], 
               y = train_data_sig$stage,
               method = "rf",
               ntree = 1000,
               tuneGrid = data.frame(mtry = mtryValues),
               preProc = c("center", "scale", "pca"),
               importance = TRUE,
               metric = "ROC",
               trControl = ctrl5f)
Sys.time() - then

rf_pre2_5cv_pred <- predict(rf_pre2_5cv, 
                            newdata = test_data_sig[, - c(1:2)])
rf_pre2_5cv_prob <- predict(rf_pre2_5cv, 
                            newdata = test_data_sig[, - c(1:2)],
                            type = "prob")

rf_pre2_5cv.df <- data.frame(stage = test_data_sig$stage, 
                             rf_pre2_5cv_pred, 
                             early_prob = rf_pre2_5cv_prob$early,
                             late_prob = rf_pre2_5cv_prob$late)

head(rf_pre2_5cv.df)

confusionMatrix(data = rf_pre2_5cv.df$rf_pre2_5cv_pred,
                reference = test_data_sig$stage,
                positive = "late")

rf_pre2_5cv_rocCurve <- roc(response = test_data_sig$stage,
                predictor = rf_pre2_5cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_sig$stage)))

auc(rf_pre2_5cv_rocCurve)

ci.roc(rf_pre2_5cv_rocCurve)

plot(rf_pre2_5cv_rocCurve, legacy.axes = TRUE)

save(rf_pre2_5cv, rf_pre2_5cv_pred, rf_pre2_5cv_prob,
     file = "../Results/rf_pre2_5cv.RData")
```

### RF/PreProc1/10-foldCV

Random forest, preprocessing method 1, 10 fold cross validataion.

```{r, message=FALSE}
then <- Sys.time()
set.seed(477)
rf_pre1_10cv <- train(x = train_data_filtered[, -c(1:2)], 
               y = train_data_filtered[, "stage"],
               method = "rf",
               ntree = 1000,
               tuneGrid = data.frame(mtry = mtryValues),
               importance = TRUE,
               metric = "ROC",
               trControl = ctrl10f)
Sys.time() - then

rf_pre1_10cv_pred <- predict(rf_pre1_10cv, 
                            newdata = test_data_filtered[, - c(1:2)])
rf_pre1_10cv_prob <- predict(rf_pre1_10cv, 
                            newdata = test_data_filtered[, - c(1:2)],
                            type = "prob")

rf_pre1_10cv.df <- data.frame(stage = test_data_filtered[, "stage"], 
                             rf_pre1_10cv_pred, 
                             early_prob = rf_pre1_10cv_prob$early,
                             late_prob = rf_pre1_10cv_prob$late)

head(rf_pre1_10cv.df)

confusionMatrix(data = rf_pre1_10cv.df$rf_pre1_10cv_pred,
                reference = test_data_filtered[, "stage"],
                positive = "late")

rf_pre1_10cv_rocCurve <- roc(response = test_data_filtered[, "stage"],
                predictor = rf_pre1_10cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_filtered[, "stage"])))

auc(rf_pre1_10cv_rocCurve)

ci.roc(rf_pre1_10cv_rocCurve)

plot(rf_pre1_10cv_rocCurve, legacy.axes = TRUE)

save(rf_pre1_10cv, rf_pre1_10cv_pred, rf_pre1_10cv_prob,
     file = "../Results/rf_pre1_10cv.RData")
```

### RF/PreProc2/10-foldCV

Random forest, preprocessing method, 10 fold cross validation

```{r, message=FALSE}
then <- Sys.time()
set.seed(477)
rf_pre2_10cv <- train(x = train_data_sig[, -c(1:2)], 
               y = train_data_sig[, "stage"],
               method = "rf",
               ntree = 1000,
               tuneGrid = data.frame(mtry = mtryValues),
               preProc = c("center", "scale", "pca"),
               importance = TRUE,
               metric = "ROC",
               trControl = ctrl10f)
Sys.time() - then

rf_pre2_10cv_pred <- predict(rf_pre2_10cv, 
                            newdata = test_data_sig[, - c(1:2)])
rf_pre2_10cv_prob <- predict(rf_pre2_10cv, 
                            newdata = test_data_sig[, - c(1:2)],
                            type = "prob")

rf_pre2_10cv.df <- data.frame(stage = test_data_sig[, "stage"], 
                             rf_pre2_10cv_pred, 
                             early_prob = rf_pre2_10cv_prob$early,
                             late_prob = rf_pre2_10cv_prob$late)

head(rf_pre2_10cv.df)

confusionMatrix(data = rf_pre2_10cv.df$rf_pre2_10cv_pred,
                reference = test_data_sig[, "stage"],
                positive = "late")

rf_pre2_10cv_rocCurve <- roc(response = test_data_sig[, "stage"],
                predictor = rf_pre2_10cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_sig[, "stage"])))

auc(rf_pre2_10cv_rocCurve)

ci.roc(rf_pre2_10cv_rocCurve)

plot(rf_pre2_10cv_rocCurve, legacy.axes = TRUE)

save(rf_pre2_10cv, rf_pre2_10cv_pred, rf_pre2_10cv_prob,
     file = "../Results/rf_pre2_10cv.RData")

```

## Naive Bayes (NB)

The last algorithm is naive Bayes.

### NB/PreProc1/5-foldCV

Naive Bayes, preprocessing method 1, 5 fold cross validation.

```{r, message=FALSE, warning=FALSE}
then <- Sys.time()
set.seed(477)
nb_pre1_5cv <- train(x = train_data_filtered[, -c(1:2)], 
               y = train_data_filtered$stage,
               method = "nb",
               ntree = 1000,
               tuneGrid = data.frame(usekernel = c(TRUE, FALSE), fL = 2),
               importance = TRUE,
               metric = "ROC",
               trControl = ctrl5f)
Sys.time() - then

nb_pre1_5cv_pred <- predict(nb_pre1_5cv, 
                            newdata = test_data_filtered[, - c(1:2)])
nb_pre1_5cv_prob <- predict(nb_pre1_5cv, 
                            newdata = test_data_filtered[, - c(1:2)],
                            type = "prob")

nb_pre1_5cv.df <- data.frame(stage = test_data_filtered$stage, 
                             nb_pre1_5cv_pred, 
                             early_prob = nb_pre1_5cv_prob$early,
                             late_prob = nb_pre1_5cv_prob$late)

head(nb_pre1_5cv.df)

confusionMatrix(data = nb_pre1_5cv.df$nb_pre1_5cv_pred,
                reference = test_data_filtered$stage,
                positive = "late")

nb_pre1_5cv_rocCurve <- roc(response = test_data_filtered$stage,
                predictor = nb_pre1_5cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_filtered$stage)))

auc(nb_pre1_5cv_rocCurve)

ci.roc(nb_pre1_5cv_rocCurve)

plot(nb_pre1_5cv_rocCurve, legacy.axes = TRUE)

save(nb_pre1_5cv, nb_pre1_5cv_pred, nb_pre1_5cv_prob,
     file = "../Results/nb_pre1_5cv.RData")
```

### NB/PreProc2/5-foldCV

Naive Bayes, preprocessing method 2, 5 fold cross validation.

```{r, message=FALSE, warning=FALSE}
then <- Sys.time()
set.seed(477)
nb_pre2_5cv <- train(x = train_data_sig[, -c(1:2)], 
               y = train_data_sig$stage,
               method = "nb",
               ntree = 1000,
               tuneGrid = data.frame(usekernel = c(TRUE, FALSE), fL = 2),
               preProc = c("center", "scale", "pca"),
               importance = TRUE,
               metric = "ROC",
               trControl = ctrl5f)
Sys.time() - then

nb_pre2_5cv_pred <- predict(nb_pre2_5cv, 
                            newdata = test_data_sig[, - c(1:2)])
nb_pre2_5cv_prob <- predict(nb_pre2_5cv, 
                            newdata = test_data_sig[, - c(1:2)],
                            type = "prob")

nb_pre2_5cv.df <- data.frame(stage = test_data_sig$stage, 
                             nb_pre2_5cv_pred, 
                             early_prob = nb_pre2_5cv_prob$early,
                             late_prob = nb_pre2_5cv_prob$late)

head(nb_pre2_5cv.df)

confusionMatrix(data = nb_pre2_5cv.df$nb_pre2_5cv_pred,
                reference = test_data_sig$stage,
                positive = "late")

nb_pre2_5cv_rocCurve <- roc(response = test_data_sig$stage,
                predictor = nb_pre2_5cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_sig$stage)))

auc(nb_pre2_5cv_rocCurve)

ci.roc(nb_pre2_5cv_rocCurve)

plot(nb_pre2_5cv_rocCurve, legacy.axes = TRUE)

save(nb_pre2_5cv, nb_pre2_5cv_pred, nb_pre2_5cv_prob,
     file = "../Results/nb_pre2_5cv.RData")
```

### NB/PreProc1/10-foldCV

Naive Bayes, preprocessing method 1, 10 fold cross validation.

```{r, message=FALSE, warning=FALSE}
then <- Sys.time()
set.seed(477)
nb_pre1_10cv <- train(x = train_data_filtered[, -c(1:2)], 
               y = train_data_filtered[, "stage"],
               method = "nb",
               ntree = 1000,
               tuneGrid = data.frame(usekernel = c(TRUE, FALSE), fL = 2),
               importance = TRUE,
               metric = "ROC",
               trControl = ctrl10f)
Sys.time() - then

nb_pre1_10cv_pred <- predict(nb_pre1_10cv, 
                            newdata = test_data_filtered[, - c(1:2)])
nb_pre1_10cv_prob <- predict(nb_pre1_10cv, 
                            newdata = test_data_filtered[, - c(1:2)],
                            type = "prob")

nb_pre1_10cv.df <- data.frame(stage = test_data_filtered[, "stage"], 
                             nb_pre1_10cv_pred, 
                             early_prob = nb_pre1_10cv_prob$early,
                             late_prob = nb_pre1_10cv_prob$late)

head(nb_pre1_10cv.df)

confusionMatrix(data = nb_pre1_10cv.df$nb_pre1_10cv_pred,
                reference = test_data_filtered[, "stage"],
                positive = "late")

nb_pre1_10cv_rocCurve <- roc(response = test_data_filtered[, "stage"],
                predictor = nb_pre1_10cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_filtered[, "stage"])))

auc(nb_pre1_10cv_rocCurve)

ci.roc(nb_pre1_10cv_rocCurve)

plot(nb_pre1_10cv_rocCurve, legacy.axes = TRUE)

save(nb_pre1_10cv, nb_pre1_10cv_pred, nb_pre1_10cv_prob,
     file = "../Results/nb_pre1_10cv.RData")
```

### nb/PreProc2/10-foldCV

Random forest, preprocessing method, 10 fold cross validation.

```{r, message=FALSE, warning=FALSE}
then <- Sys.time()
set.seed(477)
nb_pre2_10cv <- train(x = train_data_sig[, -c(1:2)], 
               y = train_data_sig[, "stage"],
               method = "nb",
               ntree = 1000,
               tuneGrid = data.frame(usekernel = c(TRUE, FALSE), fL = 2),
               preProc = c("center", "scale", "pca"),
               importance = TRUE,
               metric = "ROC",
               trControl = ctrl10f)
Sys.time() - then

nb_pre2_10cv_pred <- predict(nb_pre2_10cv, 
                            newdata = test_data_sig[, - c(1:2)])
nb_pre2_10cv_prob <- predict(nb_pre2_10cv, 
                            newdata = test_data_sig[, - c(1:2)],
                            type = "prob")

nb_pre2_10cv.df <- data.frame(stage = test_data_sig[, "stage"], 
                             nb_pre2_10cv_pred, 
                             early_prob = nb_pre2_10cv_prob$early,
                             late_prob = nb_pre2_10cv_prob$late)

head(nb_pre2_10cv.df)

confusionMatrix(data = nb_pre2_10cv.df$nb_pre2_10cv_pred,
                reference = test_data_sig[, "stage"],
                positive = "late")

nb_pre2_10cv_rocCurve <- roc(response = test_data_sig[, "stage"],
                predictor = nb_pre2_10cv.df$early_prob,
                ## This function assumes that the second
                ## class is the event of interest, so we
                ## reverse the labels
                levels = rev(levels(test_data_sig[, "stage"])))

auc(nb_pre2_10cv_rocCurve)

ci.roc(nb_pre2_10cv_rocCurve)

plot(nb_pre2_10cv_rocCurve, legacy.axes = TRUE)

save(nb_pre2_10cv, nb_pre2_10cv_pred, nb_pre2_10cv_prob,
     file = "../Results/nb_pre2_10cv.RData")

```

## Session Information

```{r}
print(sessionInfo(), locale = FALSE)
```
